Student Log-Data from a Randomized Evaluation of Educational Technology: A Causal Case Study

### 2020-05-27 paper.Rnw
issue #6, more-concrete advice adjudicating btw methods

% \begin{table}
% \begin{tabular}{p{1in}p{1.5}p{2in}p{2in}}
% & Observational Study&Mediation Analysis &Principal Stratification\\
% \hline
% Research question & How does $H$ affect $Y$& What is the role of $Z$'s
%                                              effect on $H$ in its
%                                              effect on $Y$? & Does
%                                                               $Z$'s
%                                                               effect
%                                                               on $H$
%                                                               predict its
%                                                               effect
%                                                               on $Y$?\\

>>>>

In summary: in situations where the average treatment effect of an
implementation variable can be estimated with little confounding bias
(due to the availability of high-quality covariates, good reason to
believe in random, or haphazard, variation in implementation, or other
factors), there is much to recommend the observational study approach.
A wide variety of estimation techniques are available---such as the propensity score
matching illustrated here---many of which are straightforward to carry
out and highly transparent.
The result---the ATE of implementation on the outcome---is an
intuitive estimand.

Mediation analysis can complement the observational study if,
additionally, one well-represented level of the implementation
variable is interpretable in the control condition.
For instance, in our example, $H=0$ included both low hint usage in
the treatment group and no access to hints in the control group.
In that case, mediation analysis can contextualize the estimated
effect in the observational study in terms of the overall treatment
effect, so that implementation can be interpreted as a causal
mechanism.

However, if unobserved confounding is a serious threat, and complex
statistical modeling (and model checking) is not a practical barrier,
PS may be a better choice.
% \begin{table}
% \begin{tabular}{p{1in}p{1.5}p{2in}p{2in}}
% & Observational Study&Mediation Analysis &Principal Stratification\\
% \hline
% Research question & How does $H$ affect $Y$& What is the role of $Z$'s
%                                              effect on $H$ in its
%                                              effect on $Y$? & Does
%                                                               $Z$'s
%                                                               effect
%                                                               on $H$
%                                                               predict its
%                                                               effect
%                                                               on $Y$?\\
